---
title: "Homework 24 Scratch version"
output: html_notebook
---

# Overview
Today we are going to jump back into the GSWR book and tackle the first half of chapter 5. Let's do it.

```{r}
# First set up the environment to start with a clean slate and the libraries we will need
rm(list = ls())

library(ggplot2)
library(dplyr)
library(here)
```
Now I am going to upload the data that we will be using to work thorugh this dataset. 

```{r}
# import the csv file that we are going to use for this exercise without stringsAsFactors = TRUE
lady <- read.csv(here("Data", "ladybirds.csv"), stringsAsFactors = TRUE)

# now we are going to run a summary of Leaves to take a look at the data
glimpse(lady)
```
Ok, now that the dataset is in R, we are going to take a look at the first type of test that the book covers - chi squared contingency table analysis. 

# Chi squared contingency table analysis
The chi squared test is an analysis of count data.  As such, it measures if there is any association between any number of categorical variables (obviously at least greater than 1). 
For the ladybird data, the purpose of performing the chi squared analysis is to test the null hypothesis that there is no association between ladybird color and the habitat they live in. Essentially, it is to see if there is a biological advantage for certain lady birds to live in different habitats based on their colors. 

Earlier, I used the `glimpse()` function to take a look at the data.  There are a few different variables that we are most interested in such as "Habitat", "colour", and "number".  Before we perform the analysis, we want to organize the data to prime it for plotting and analysis. This can be done using the `group_by()` and `summarize()` functions.
```{r}
# use piping to summarize the data using the group_by and summarize funcitons
totals <- lady %>% 
     group_by(Habitat, colour)%>%
     summarize(total.number = sum(number))

# make a bar graph of this data to get a quick overview of it
ggplot(totals, aes(x = Habitat, y = total.number, fill = colour))+

# tell R not to calculate anything from the data and place the bars next to each other
     geom_bar(stat = 'identity', position = 'dodge') +
     
# manually define the colors of the plot
     scale_fill_manual(values = c(black = "black", red = "red"))

```
The purpose of this bar graph is to show the data visually.  In this case we are not summarizing anything so the bar chart is acceptable. Here "fill = morph_colour" is added to the `aes()` function to fill in the bars with colour.  

Also the function `geom_bar()` has the following parameters:
- stat = 'identity' is used to tell ggplot2 not to calculate anything
- position = 'dodge' tells ggplot2 to put the bars in each Habitat group next to each other

Based on this data, it appears the black morph is more common in the industrial area compared to the red morph.  The opposite relationship appears to be true for the rural area.  The chi squared test compares frequencies and this graph makes it seem as if morphs are not equally distriuted among the habitats.  Therefore we are looking to reject the null hypothesis.

```{r}
# give an overview of the data
totals

# transform the data into a matrix form that will be used for the chi squared analysis
lady.mat <- xtabs(number ~ Habitat + colour, data = lady)

lady.mat

# perform the chi squared analysis using the matrix above
chisq.test(lady.mat)
```

Now we can actually do something with this analysis.  Since our P value is very small, it suggests that there is a level of significance to the diferences between the color of morphs found in rural and industrial areas.



