---
title: "Practicing Data Modeling with Soil Data"
output: html_notebook
---

For this practice set, I will be taking soil data that was given to us in class and using that to create some different plots to model the data. The first step will be to clear R's environment and import the libraries that we will need.

```{r}
# clear the R environment and import the libraries that we need
rm(list = ls())
library(here)
library(ggplot2)
library(dplyr)
```

The next thing to do is to import the data.  The `here()` function sets the file path to the homework folder that I am currently working in.  Using `read.csv()` we can select the soil data from the "Data" folder in this project. 

```{r}
# import the data 
soil <- read.csv(here("Data", "soil_data.csv"), stringsAsFactors = TRUE)

# open the data to ensure that it imported correctly
View(soil)
```

Now, we want to create a scatterplot, boxplot, and histogram for the SOM_percent column. The first thing I am going to do however is use the `filter()` function to take all of the ones with DEGRASS in the name and `mutate()` to create a new column called DEGRASS.  This process will be repeated with all 5 forests.

```{r}
# use piping to filter out the Degrasse Plots from the rest of the forest data name and make those samples into a new dataframe
Degrasse <- c('DEGRASSE PLOT 1 SAMPLE 1', 'DEGRASSE PLOT 1 SAMPLE 2', 'DEGRASSE PLOT 1 SAMPLE 3', 'DEGRASSE PLOT 2 SAMPLE 1','DEGRASSE PLOT 2 SAMPLE 2', 'DEGRASSE PLOT 2 SAMPLE 3', 'DEGRASSE PLOT 3 SAMPLE 1', 'DEGRASSE PLOT 3 SAMPLE 2', 'DEGRASSE PLOT 3 SAMPLE 3')

Donnerville <- c('DONNERVILLE PLOT 1 SAMPLE 1', 'DONNERVILLE PLOT 1 SAMPLE 2', 'DONNERVILLE PLOT 1 SAMPLE 3', 'DONNERVILLE PLOT 2 SAMPLE 1','DONNERVILLE PLOT 2 SAMPLE 2', 'DONNERVILLE PLOT 2 SAMPLE 3', 'DONNERVILLE PLOT 3 SAMPLE 1', 'DONNERVILLE PLOT 3 SAMPLE 2', 'DONNERVILLE PLOT 3 SAMPLE 3')

Kip <- c('KIP TRACT PLOT 1 SAMPLE 1', 'KIP TRACT PLOT 1 SAMPLE 2', 'KIP TRACT PLOT 1 SAMPLE 3', 'KIP TRACT PLOT 2 SAMPLE 1','KIP TRACT PLOT 2 SAMPLE 2', 'KIP TRACT PLOT 2 SAMPLE 3', 'KIP TRACT PLOT 3 SAMPLE 1', 'KIP TRACT PLOT 3 SAMPLE 2', 'KIP TRACT PLOT 3 SAMPLE 3')

Peavine <- c('PEAVINE PLOT 1 SAMPLE 1', 'PEAVINE PLOT 1 SAMPLE 2', 'PEAVINE PLOT 1 SAMPLE 3', 'PEAVINE PLOT 2 SAMPLE 1','PEAVINE PLOT 2 SAMPLE 2', 'PEAVINE PLOT 2 SAMPLE 3', 'PEAVINE PLOT 3 SAMPLE 1', 'PEAVINE PLOT 3 SAMPLE 2', 'PEAVINE PLOT 3 SAMPLE 3')

Hammond <- c('HAMMOND PLOT 1 SAMPLE 1', 'HAMMOND PLOT 1 SAMPLE 2', 'HAMMOND PLOT 1 SAMPLE 3', 'HAMMOND PLOT 2 SAMPLE 1','HAMMOND PLOT 2 SAMPLE 2', 'HAMMOND PLOT 2 SAMPLE 3', 'HAMMOND PLOT 3 SAMPLE 1', 'HAMMOND PLOT 3 SAMPLE 2', 'HAMMOND PLOT 3 SAMPLE 3')

soil %>% filter(Field.Name %in% Degrasse) %>% mutate(Forest =  )
```

I decided to filter out the Degrasse forest data and put it in a new data frame becuause I ran into issues when I tried to use the `mutate()` function to add these in a seperate column to the soil DF.  R was having trouble becuase the Degrasse DF only has 9 observations whereas the soil DF has 45 observations.  Filtering out the Degrasse data and putting it into a new DF should allow me to do all of the same functions and complete the same goals of this assignment.
