---
title: "Practicing Data Modeling with Soil Data"
output: html_notebook
---

For this practice set, I will be taking soil data that was given to us in class and using that to create some different plots to model the data. The first step will be to clear R's environment and import the libraries that we will need.

```{r}
# clear the R environment and import the libraries that we need
rm(list = ls())
library(here)
library(ggplot2)
library(dplyr)
```

The next thing to do is to import the data.  The `here()` function sets the file path to the homework folder that I am currently working in.  Using `read.csv()` we can select the soil data from the "Data" folder in this project. 

```{r}
# import the data 
soil <- read.csv(here("Data", "soil_data.csv"), stringsAsFactors = TRUE)

# open the data to ensure that it imported correctly
View(soil)
```

Now, we want to create a scatterplot, boxplot, and histogram for the SOM_percent column. The first thing I am going to do however is use the `filter()` function to take all of the ones with DEGRASS in the name and `mutate()` to create a new column called DEGRASS.  This process will be repeated with all 5 forests.

```{r}
# use piping to filter out the Degrasse Plots from the rest of the forest data name and make those samples into new columns
Degrasse_plts <- c('DEGRASSE PLOT 1 SAMPLE 1', 'DEGRASSE PLOT 1 SAMPLE 2', 'DEGRASSE PLOT 1 SAMPLE 3', 'DEGRASSE PLOT 2 SAMPLE 1','DEGRASSE PLOT 2 SAMPLE 2', 'DEGRASSE PLOT 2 SAMPLE 3', 'DEGRASSE PLOT 3 SAMPLE 1', 'DEGRASSE PLOT 3 SAMPLE 2', 'DEGRASSE PLOT 3 SAMPLE 3')
d <- soil %>% filter(Field.Name %in% Degrasse_plts)
View(d)
```

