---
title: "Practicing using R to perform different statistical analyses"
output: html_notebook
---

# Overview
This homework is an extension of the work we did in class on analysing the soil data set.  Here are the questions we are trying to answer:

1. Is there a relationship between Soil organic matter (SOM, %) and soil calcium (Ca_ppm)?

2. Does the abundance of worms differ between Donnerville and the Kip Tract?

3. Does the abundance of soil Magnesium (Mg_ppm) differ between Peavine and South Hammond?

4. Does the abundance of zinc (Zn_ppm) differ among forests?

First let's set things up:
```{r}
rm(list =ls())
library(tidyverse)
library(here)
library(ggfortify)
```

Now lets pull the data in:
```{r}
# import the dataset
soil <- read.csv(here("Data", "soil_reduced-1.csv"), stringsAsFactors = TRUE)

View(soil)
```

# Checking the data
Before we proceed with the analysis, it is important to check the data for possible errors and/or outliers. 

A. Forest names
```{r}
# checking the levels of forests to make sure the names are imported correctly
levels(soil$Forest)
```
B. Soil organic matter and soil calcium
```{r}
# make a histogram to check and see if there are any outliers
ggplot(soil, aes(x = SOM_percent))+
     geom_histogram()

```
Based on this histogram, there are not really any distict outliers.  Some of the values are high but these data were recorded from a machine so it is unlikely that the higher values are true outliers.

```{r}
# checking the soil calcium with a histogram
ggplot(soil, aes(x = Ca_ppm))+
     geom_histogram()
```
These data points also look good.  Let's run a summary to get an brief overview of the spread of the data.

```{r}
# use the summary to get some basic stats on the data set
summary(soil)
```
Of these variables, the Buffer pH, Nitrate levels, and worms appear to have some potential outliers.  for these, the max is very far from the mean which indicates there may have been something that was a little off. 

## 1. Is there a relationship between soil organic matter (SOM, %) and soil calcium(Ca_ppm)?
Both variables are continuous so lets take a look at these with a scatterplot.

### Plot the relationship
```{r}
# make a scatterplot to take a look at how these are related
ggplot(soil, aes(x = SOM_percent, y = Ca_ppm, colour = Forest))+
     geom_point()+
     theme_bw()
```
From looking at the data, it appears there may be a weak positive relationship between SOM and Ca. The influence of a single point on the relationship is called "leverage".  We will be taking a look more into that by trying to figure out if there is a linear relationship between these two variables. 

Can I guess the slope or y-intercept? 
- the rise is about 1000
- the run is about 53
Dividing these to estimate the slope would give us (1000/53) = 18 ish
y-intercept is about -5 or so.


What statistical model fits these data?
Both the predictor and the response variables are continuous so we will try to model this with a linear regression. Let's construct the model
```{r}
# model the relationship between these two variables with a linear regression, for the formula argument, you are saying "y as a function of x"
SOM_CA_mod <- lm(Ca_ppm ~ SOM_percent, data = soil)
```

Now we need to check the assumptions before we move forward with the model.  We will do this by using the `autoplot()` function out of the `ggfortify` libarary.

```{r}
# check the assumptions
autoplot(SOM_CA_mod)
```
From the looks of the Residuals vs Fitted plot, it appears there is a lot of variance at the start of the fitted values.  This makes sense because a lot of the data points are clustered at the begining and there are a decent amount of higher valeus within those data points.  That being said, we may be violating this assumption because the residuals are so far from the fitted data. 

Based on these observations however and since these data came from a lab, I will say that these data do fit the assumption criteria of the linear model.

Now let's look at the model outcome
```{r}
# first check this with ANOVA(), thid does not calculate an ANOVA but it provides the standard output of the ANOVA
anova(SOM_CA_mod)
```
```{r}
# run the summary of the function
summary(SOM_CA_mod)
```

In this case, the null hypothesis is that there is no meaningful relationship between soil organic matter and calcium ppm.  Since the p value is greater than 0.05, it indicates there is no significant relationship between these two variables and we accept the null hypothesis. 


## 2. Does the abundance of worms differ between Donnerville and the Kip Tract?
The first thing that we are going to do is plot these variables.  Since we are looking at the abundance of worms between donnerville and the kip tract, we will plot these as a histogram
```{r}
# isolate the Donnerville and Kip Tract data
worms <- soil %>% filter(Forest == c("KIP TRACT", "DONNERVILLE"))
# plot histogram modeling worms between Kip tract and Donnerville
ggplot(worms, aes(x = Worms))+
    geom_histogram(binwidth = 1)+
     facet_wrap(~Forest)
```
From this histogram, the Kip Tract and Donnerville locations, it appears that there may be a higher abundance of worms in the Kip Tract compares to Donnerville. For this example, the worm data is quantitative ratio because it is counting data. On the other hand, the location is categorical nominal data since there are a number of differnet locations.  Since this is a count data set, we will analyse it using a chi squared test.

```{r}
# calculate the means of the ozone levels in each location.
totals <- worms %>% 
     group_by(Forest)%>%
     summarize(total_worms = sum(Worms))
```
Another way to take a look at this relationship is through making a bar graph of the total worms.

```{r}
ggplot(totals, aes(x = Forest, y = total_worms))+
    geom_bar(stat = "identity", position = "dodge")
```


This gives us some more insight into the abundance of worms between the Kip Tract and Donnerville.  The means appear to be a bit further away from each other but it is still worth taking a closer look at with the chi sqared test.

```{r}
# transform the data into a matrix form that will be used for the chi squared analysis
worms.mat <- xtabs(Worms ~Forest , data = worms)

worms.mat

# perform the chi squared analysis using the matrix above
chisq.test(worms.mat)

# take a closer look at the stats generated by the chi squared analysis
worms.mat <- chisq.test(worms.mat)
names(worms.mat)
```
Based on the chi squared test, we reject the null hypothesis because our p-value is less than 0.05 which indicates there is a significantly higher abundance of worms at the Kip Tract compared to Donnerville.

## 3. Does the abundance of soil Magnesium (Mg_ppm) differ between Peavine and South Hammond?
For this data analysis, I will first graph the abundance of magnesium in Peavine and compare it to south Hammond using a set of histograms.
```{r}
# isolate the Peavine and South Hammond data
mag <- soil %>% filter(Forest == c("PEAVINE", "SOUTH HAMMOND"))
# plot histogram modeling soil magnesium between Peavine and South Hammond
ggplot(mag, aes(x = Mg_ppm))+
    geom_histogram(binwidth = 1)+
     facet_wrap(~Forest)
```
This histogram gives us some insight into the relative amount of magnesium in the soil samples from each of the plots. Upon first glance it looks like the data for the South Hammond area is skewed to the left - meaning it may have a greater amount of soil magnesium compared to the Peavine area.  

We can take a closer look at this statistically however using a two sample t test.  The two forest locations are categorical nominal data and the mg ppm data is quantitative ratio.  Because of this, it make sense to take a look at these using a two sample T test.

```{r}
# calculate the means of the magnesium levels in each location.
mag %>% 
     group_by(Forest)%>%
     summarize(mean_magppm = mean(Mg_ppm))
```
Now we can use this data to perform a two sample t test.
```{r}
# perform the two sample t test of the data
t.test(Mg_ppm ~ Forest, data = mag)
```
Though the histogram makes it look like there may be a difference between the magnesium within the soil at the Peavine location and the South Hammond locations, the two sample t test says otherwise.  Since the p-value is greater than 0.05, we accept the null hypothesis which states that there is no significant difference between the magnesium concentration in the Peavine vs South Hammond locations. 


## 4. Does the abundance of zinc (Zn_ppm) differ among forests?
The last think that we are going to look at is the zinc abundace.  Since we are looking at this between the different forests, we will first plot the zinc abundance as a series of histograms grouped by forest. 
```{r}
# plot histogram modeling soil zinc between the forests
ggplot(soil, aes(x = Zn_ppm))+
    geom_histogram(binwidth = 1)+
     facet_wrap(~Forest)
```
From this, there are some clear differences in the distribution of zinc abundance between the forests.  To see whether this is significant or not, we will perform an ANOVA analysis.  This test makes the most sense in this context because we are trying to compare the zinc abundance (quantitative ratio data) with multiple Forests (categorical nominal data).

```{r}
# Perform the ANOVA
model_zinc <- lm(Zn_ppm ~ Forest, data = soil)

# check the assumptions
autoplot(model_zinc, smooth.colour = NA)

summary(model_zinc)
```
The results of the ANOVA show that there is a significant difference in zinc abundance between the Degrasse location and the others.  Between the rest of the forests, there is no significant difference in zinc soil abundance.  Let's replot this to see the differences a little better:
```{r}
# modeling zinc abundance, and the differences between the forests and parasite manipulations
ggplot(soil, aes(x = Forest, y = Zn_ppm, colour = Forest)) +
        geom_point() +
        coord_flip()+
        theme_bw()
```









