---
title: "Homework 12 - Practicing Manipulating Data"
output: html_notebook
---

For this homework assignment, we will be using a new dataset to practice a number of different functions within the `dplyr()` library.  Before we do anything though, we need to clear R's environment window and import the appropriate libraries so we can acess these functions. 

```{r}
# clearing the environment window
rm(list = ls())

# importing the appropriate libraries
library(dplyr)
library(ggplot2)
library(here)
```
Now we can import the dataset that we want to use for this exercise. I used the `here()` function to start defining the file path so we can use the `read.csv()` function without writing out the entire file path.  This process goes as follows:

```{r}
# importing the data set and assigning it to a variable
cover = read.csv(here("Data", "donnerville_percent_cover.csv"), stringsAsFactors = TRUE)
```

Now we will check the column headings for the dataset and extract some information about the types of data within it using the `head()`, `glimpse()` and `str()` functions.

```{r}
# check the column headings of the data and the first few rows of data
head(cover)

# provides basic summary of data
glimpse(cover)

# use the str() function to take a look at the data structure
str(cover)
```
Let's rename the "Plot.dim" and "Plot.type" column headings to see how this works.

```{r}
# use piping to reassign the names of the data
cover <- cover %>% rename(Plot_type = Plot.type, Plot_dim = Plot.dim)

# check the names to see if it changed. 
names(cover)

# another way to do this is to use the output of the names function and reassign them
names(cover) <- c("Site", "Date", "Plot_type", "Plot_dim", "Slope", "Aspect", "Plot", "Type", "Percent")
```

Now the goal is to determine the average percent cover for Donnerville 1 by the type of cover (bare, bryophytes, vascular plants).  To do this, I will have to select the rows for the Donnerville 1 plots and then filter the types of cover. Once I filter by types of cover, I can then find the average percent cover for each type.

```{r}
# get just the plot 1 data
plt1 <- filter(cover, Site == "Donnerville 1")
```

```{r}
# use the group_by() function to determine the average percent cover for each type of leaf cover
plt1 %>% group_by(Type) %>% summarize(avg_percent_cover = mean(Percent))

```
Now that we have the mean percent cover for each type for the plot 1 data, we can move onto the next task.  The next task is to determine the average percent vascular plant cover across all sites.  To do this, I will first sort the data and select the types with vascular plant cover using the `filter()` function

```{r}
# isolate the vascular plant cover data
vascular <- filter(cover, Type == "Vascular")
```

Now, I will use the `mean()` function to determine the average percent cover for the vascular type of plant cover

```{r}
vascular %>% summarize(avg_percent = mean(Percent))
```

