---
title: "Homework 27 - Getting Familliar With Data Visualization in R "
output: html_notebook
---

# Overview
For this homework assignment, I will be working through the first and second chapters of the "ggplot2: Elegant Graphics for Data Analysis". The goal of this activity is to learn a little bit more about how we can use the different functions inside of the `ggplot2` package to make visually appealing and clear figures with our data. 

The first thing that I will do is set up the environement by running the code to install the different libraries that we need:

```{r}
# clear the environment window
rm(list=ls())

# import the libraries that we will need to use
library(ggplot2)
library(tidyverse)
```
Now we are going to use the MPG data for this practice.  The MPG data is built into the ggplot frame

```{r}
# taking a look at the dataset
mpg
```
## Taking a look at the key components of a ggplot2 plot

Every plot created in ggplot2 has three main parts: the data, the aesthetic mappings, and the geom function

```{r}
# example of the ggplot functionality
ggplot(mpg, aes(x = displ, y = hwy)) + 
  geom_point()

# this can also be written as:
# ggplot(mpg, aes(displ, hwy)) +
  # geom_point()
```
The data in this case is the mpg data, the aes function refers to the engine size mapped to the x position and the fuel economy to the y position.  The layer is the points to create a scatterplot.

Describe the data, aesthetic mappings, and layers for the following plots:
```{r}
ggplot(mpg, aes(cty, hwy)) + geom_point()

ggplot(diamonds, aes(carat, price)) + geom_point()

ggplot(economics, aes(date, unemploy)) + geom_line()

ggplot(mpg, aes(cty)) + geom_histogram()
```
1. Data = mpg, aes = cty as the x and hwy as the y, geom = point to create a scatterplot

2. Data = diamonds, aes = carat as the x and price as the y, geom = point to create a scatterplot

3. Data = economics, aes = date as the x and unemployment as the y, aes = line to show a line graph modeling the trend

4. Data = mpg, aes = cty, geom = histogram to shot the spred of the city data

## Color, size, shape, and other aesthetic attributes
There are a number of different ways to change and customize your plot from within the `aes()` function.  Some examples include:

```{r}
aes(displ, hwy, colour = class)
aes(displ, hwy, shape = drv)
aes(displ, hwy, size = cyl)
```
We can overlay these variables in the previous scatterplots by adding them directly in the `aes()` function:

```{r}
ggplot(mpg, aes(displ, hwy, colour = class)) +
  geom_point()
```
Here each point now has a unique color that represents its class.  This also puts a legend on the graph that makes it easier to interpret.

You can also add these features by adding another layer to the plot:
```{r}
ggplot(mpg, aes(displ, hwy)) + geom_point(aes(colour = "blue"))

ggplot(mpg, aes(displ, hwy)) + geom_point(colour = "blue")
```
When you use the `aes()` function inside of a new layer, it adds a legend but the color is incorrect.  If you do not use the `aes()` function, the color will be correct but no legend will be generated. 

## Faceting
Another way to visually show data is to split it up by category. This allows you to split the data into subsets to make it easier to digest.  For example:
```{r}
ggplot(mpg, aes(displ, hwy)) + 
  geom_point() + 
  facet_wrap(~class)
```
The faceting is added as another layer to the plot and you can specify how you want the data to be faceted depenging on the goals of your data visualization. 

## Plot geoms
This is the primary method used to change the type of plot that you are trying to make.  Here are a few different options:
     - geom_smooth() fits a smoother to the data to show standard error
     - geom_boxplot() creates box plot
     - geom_histogram() and geom_freqpoly() shows distribution of continuous variables
     - geom_bar() shows distriution of categorical variables
     - geom_path() and geom_line() draw lines between data points
     
### Adding a smoother to a plot
One way to judge the trends in the data is to use a smoother to look at the dominant pattern.
```{r}
ggplot(mpg, aes(displ, hwy)) + 
  geom_point() + 
  geom_smooth()
#> `geom_smooth()` using method = 'loess' and formula 'y ~ x'
```
Here the smoother is overlayed ontop of a scatterplot.  You can also see the standard error in the grey highlighted portion of the graph (these are point-wise confidence intervals and can be turned off with `geom_smooth(se = FALSE)`).

For the `geom_smooth()` function, the method argument allows you to choose the type of model you want to smooth the curve.
     - method = "loess" is the default for small n.  It uses a smooth local regression and the span parameter controls the "wiggliness" of the line
```{r}
ggplot(mpg, aes(displ, hwy)) + 
  geom_point() + 
  geom_smooth(span = 0.2)
#> `geom_smooth()` using method = 'loess' and formula 'y ~ x'

ggplot(mpg, aes(displ, hwy)) + 
  geom_point() + 
  geom_smooth(span = 1)
#> `geom_smooth()` using method = 'loess' and formula 'y ~ x'
```
     - method = "gam" fits a generalized additive model that is provided by the mgcv package.  This requires the mgcv package to function and a formula like y ~ s(x)
```{r}
library(mgcv)
ggplot(mpg, aes(displ, hwy)) + 
  geom_point() + 
  geom_smooth(method = "gam", formula = y ~ s(x))
```
     - method = "lm" fits a linear model and gives the line of best fit
```{r}
ggplot(mpg, aes(displ, hwy)) + 
  geom_point() + 
  geom_smooth(method = "lm")
#> `geom_smooth()` using formula 'y ~ x'
```
     - method = "rlm" is similar to the linear regression created by the lm() function but uses a differnt fitting algorithm that discounts outliers more, this requires the MASS package. 
     
### Boxplots and jittered points
Boxplots are useful to model the spread of continuous variables by categorical levels.  One way to do this is to create a scatterplot like this:
```{r}
ggplot(mpg, aes(drv, hwy)) + 
  geom_point()
```
Say we wanted to clean this up a bit because all of the points are very much on top of each other and it is tough to see the distribution, there are a few ways we can do this. 
     - geom_jitter() adds some random noise to the data to avoid overplotting
     - geom_boxplot() sumarizes the shape of the distribution
     - geom_violin() shows the density of the distribution
     
```{r}
ggplot(mpg, aes(drv, hwy)) + geom_jitter()
ggplot(mpg, aes(drv, hwy)) + geom_boxplot()
ggplot(mpg, aes(drv, hwy)) + geom_violin()
```
### Histograms and frequency polygons
Histograms and frequency pologons show the spread and distribution of data.  They do not show different summary statistics like boxplots do but they are able to show the spread really well:
```{r}
ggplot(mpg, aes(hwy)) + geom_histogram()
#> `stat_bin()` using `bins = 30`. Pick better value with `binwidth`.
ggplot(mpg, aes(hwy)) + geom_freqpoly()
#> `stat_bin()` using `bins = 30`. Pick better value with `binwidth`.
```
Both of these bin the data and count the number of observations in each bin.  You can control the width of the bins using the `binwidth` argument.  This can give you a better idea of the spread of the data, especially if it is really small. 

```{r}
ggplot(mpg, aes(hwy)) + 
  geom_freqpoly(binwidth = 2.5)
ggplot(mpg, aes(hwy)) + 
  geom_freqpoly(binwidth = 1)
```
You can also create a density plot using the `geom_density()` function but these can be a bit harder to interpret.  There are a few different ways you can compare the distributions using the frequency polygon and it even allows you to add in faceting:
```{r}
ggplot(mpg, aes(displ, colour = drv)) + 
  geom_freqpoly(binwidth = 0.5)
ggplot(mpg, aes(displ, fill = drv)) + 
  geom_histogram(binwidth = 0.5) + 
  facet_wrap(~drv, ncol = 1)
```
### Bar charts
Bar charts are best for comparing count data between categories:
```{r}
ggplot(mpg, aes(manufacturer)) + 
  geom_bar()
```
This chart consists of unsummarized data.  Each observation contributes one unit to the height of each bar.  If you have presummarized data however it is a little different:
```{r}
drugs <- data.frame(
  drug = c("a", "b", "c"),
  effect = c(4.2, 9.7, 6.1)
)

# tell geom_bar not to run the default tat which bins and counts the data
ggplot(drugs, aes(drug, effect)) + geom_bar(stat = "identity")
ggplot(drugs, aes(drug, effect)) + geom_point()
```
### Time series with line and path plots
Line and path plots are commonly used for time series data.  Line plots joing the points from left to right and the path joins them in the order that they appear in the dataset.  For example:
```{r}
ggplot(economics, aes(date, unemploy / pop)) +
  geom_line()
ggplot(economics, aes(date, uempmed)) +
  geom_line()
```
These are great for showing how things change over time.  If we want to draw both of the time series on the same plot, we can join points adjacent in time with line segements.

```{r}
ggplot(economics, aes(unemploy / pop, uempmed)) + 
  geom_path() +
  geom_point()

year <- function(x) as.POSIXlt(x)$year + 1900
ggplot(economics, aes(unemploy / pop, uempmed)) + 
  geom_path(colour = "grey50") +
  geom_point(aes(colour = year(date)))
```
From these, you can see that unepmployment rate and the length of unemployment are highly correlated but it appears that the length of unemployment is now rising in recent years.

## Modifying the axes
There are a few different ways that we can modify the axes when we are analysing data.  The method that we have been using is with the `xlab()` and `ylab()` layers.
```{r}
ggplot(mpg, aes(cty, hwy)) +
  geom_point(alpha = 1 / 3)

ggplot(mpg, aes(cty, hwy)) +
  geom_point(alpha = 1 / 3) + 
  xlab("city driving (mpg)") + 
  ylab("highway driving (mpg)")

# Remove the axis labels with NULL
ggplot(mpg, aes(cty, hwy)) +
  geom_point(alpha = 1 / 3) + 
  xlab(NULL) + 
  ylab(NULL)
```






