---
title: "Homework 27 - Getting Familliar With Data Visualization in R "
output: html_notebook
---

# Overview
For this homework assignment, I will be working through the first and second chapters of the "ggplot2: Elegant Graphics for Data Analysis". The goal of this activity is to learn a little bit more about how we can use the different functions inside of the `ggplot2` package to make visually appealing and clear figures with our data. 

The first thing that I will do is set up the environement by running the code to install the different libraries that we need:

```{r}
# clear the environment window
rm(list=ls())

# import the libraries that we will need to use
library(ggplot2)
library(tidyverse)
```
Now we are going to use the MPG data for this practice.  The MPG data is built into the ggplot frame

```{r}
# taking a look at the dataset
mpg
```
## Taking a look at the key components of a ggplot2 plot

Every plot created in ggplot2 has three main parts: the data, the aesthetic mappings, and the geom function

```{r}
# example of the ggplot functionality
ggplot(mpg, aes(x = displ, y = hwy)) + 
  geom_point()

# this can also be written as:
# ggplot(mpg, aes(displ, hwy)) +
  # geom_point()
```
The data in this case is the mpg data, the aes function refers to the engine size mapped to the x position and the fuel economy to the y position.  The layer is the points to create a scatterplot.

Describe the data, aesthetic mappings, and layers for the following plots:
```{r}
ggplot(mpg, aes(cty, hwy)) + geom_point()

ggplot(diamonds, aes(carat, price)) + geom_point()

ggplot(economics, aes(date, unemploy)) + geom_line()

ggplot(mpg, aes(cty)) + geom_histogram()
```
1. Data = mpg, aes = cty as the x and hwy as the y, geom = point to create a scatterplot

2. Data = diamonds, aes = carat as the x and price as the y, geom = point to create a scatterplot

3. Data = economics, aes = date as the x and unemployment as the y, aes = line to show a line graph modeling the trend

4. Data = mpg, aes = cty, geom = histogram to shot the spred of the city data

## Color, size, shape, and other aesthetic attributes
There are a number of different ways to change and customize your plot from within the `aes()` function.  Some examples include:

```{r}
aes(displ, hwy, colour = class)
aes(displ, hwy, shape = drv)
aes(displ, hwy, size = cyl)
```
We can overlay these variables in the previous scatterplots by adding them directly in the `aes()` function:

```{r}
ggplot(mpg, aes(displ, hwy, colour = class)) +
  geom_point()
```
Here each point now has a unique color that represents its class.  This also puts a legend on the graph that makes it easier to interpret.

You can also add these features by adding another layer to the plot:
```{r}
ggplot(mpg, aes(displ, hwy)) + geom_point(aes(colour = "blue"))

ggplot(mpg, aes(displ, hwy)) + geom_point(colour = "blue")
```
When you use the `aes()` function inside of a new layer, it adds a legend but the color is incorrect.  If you do not use the `aes()` function, the color will be correct but no legend will be generated. 

## Faceting
Another way to visually show data is to split it up by category. This allows you to split the data into subsets to make it easier to digest.  For example:
```{r}
ggplot(mpg, aes(displ, hwy)) + 
  geom_point() + 
  facet_wrap(~class)
```
The faceting is added as another layer to the plot and you can specify how you want the data to be faceted depenging on the goals of your data visualization. 





