---
title: "Homework 21 - Working With Data in R"
output: html_notebook
---

# Overview
During this homework assignment we are going to take a look at different ways that you can visualize data.  To start, I am going to import the libraries that we are going to need to import and manipulate the data. 

```{r}
# clear the R environment and import the libraries that we need
rm(list = ls())
library(here)
library(ggplot2)
library(dplyr)
```

The data that we are going to be working with is the "compensation.csv" data that we have been working with. To import it, I will use the `read.csv()` and the `here()` functions.

```{r}
# import the data 
compensation <- read.csv(here("Data", "compensation.csv"))

# open the data to ensure that it imported correctly
View(compensation)
```

# Working with `ggplot2()`
The first thing that we are going to do is try and create a simple scatter plot of the data.  We will do this using `ggplot2()`.

```{r}
# Use ggplot to make a simple scatterplot
# the function takes the following parameters: ggplot(data frame, aes(x = x-axis, y = y-axis))+geom_point()

# the +geom_point() function adds another layer to the plot that will overlay your data points
ggplot(compensation, aes(x = Root, y = Fruit)) + geom_point()
```
## Customizing plots
One thing that we notice about the scatterplot above is the grey background.  This is not always ideal and we can use the following code to get rid of the grey.

```{r}
# getting rid of the grey in the plot background
ggplot(compensation, aes(x = Root, y = Fruit)) + geom_point() + 
     theme_bw()
```

