---
title: "Homework 21 - Working With Data in R"
output: html_notebook
---

# Overview
During this homework assignment we are going to take a look at different ways that you can visualize data.  To start, I am going to import the libraries that we are going to need to import and manipulate the data. 

```{r}
# clear the R environment and import the libraries that we need
rm(list = ls())
library(here)
library(ggplot2)
library(dplyr)
```

The data that we are going to be working with is the "compensation.csv" data that we have been working with. To import it, I will use the `read.csv()` and the `here()` functions.

```{r}
# import the data 
compensation <- read.csv(here("Data", "compensation.csv"))

# open the data to ensure that it imported correctly
View(compensation)
```

# Working with `ggplot2()`
The first thing that we are going to do is try and create a simple scatter plot of the data.  We will do this using `ggplot2()`.

```{r}
# Use ggplot to make a simple scatterplot
# the function takes the following parameters: ggplot(data frame, aes(x = x-axis, y = y-axis))+geom_point()

# the +geom_point() function adds another layer to the plot that will overlay your data points
ggplot(compensation, aes(x = Root, y = Fruit)) + geom_point()
```
## Customizing plots
One thing that we notice about the scatterplot above is the grey background.  This is not always ideal and we can use the following code to get rid of the grey.

```{r}
# getting rid of the grey in the plot background
ggplot(compensation, aes(x = Root, y = Fruit)) + geom_point() + 
     theme_bw()
```
Another thing we can adjust is increasing the size of the points.  To do this, we will add a `size` argument to the `geom_points()` layer. 

```{r}
# increasing the size of the data points
ggplot(compensation, aes(x = Root, y = Fruit)) + geom_point(size = 5) +
     theme_bw()
```
We can also customize the x and y-axis labels by using `xlab()` and `ylab()`.  This allows us to get a little more creative and specific with the labels as opposed to them defaulting to the titles of the data we are using. 

```{r}
ggplot(compensation, aes(x = Root, y = Fruit)) +
geom_point(size = 5) +
xlab("Root Biomass") +
ylab("Fruit Production") +
theme_bw()
```
The `ggplot2()` library also allows you to change the color of the data points based on the levels.  For this, you add another argument to the `aes()` function.  After specifying the x and the y, you can add a color as modeled below. 

```{r}
# differentiating between the levels of data by changing the colors of the levels
ggplot(compensation, aes(x = Root, y = Fruit, colour = Grazing)) +
geom_point(size = 5) +
xlab("Root Biomass") +
ylab("Fruit Production") +
theme_bw()
```
Another way to differentiate between different levels is to specify the shape of the data.  Here, we can do that by substituting the `colour = Grazing` in the code above with `shape = Grazing`.

```{r}
# differentiating between levels of data by changing the shape of the data points
ggplot(compensation, aes(x = Root, y = Fruit, shape = Grazing)) +
geom_point(size = 5) +
xlab("Root Biomass") +
ylab("Fruit Production") +
theme_bw()
```
## Box and Whisker Plots
Up to this point, we have primarily been working with scatterplots but the `ggplot2()` library can do a lot more than that.  To make a box plot, we will use the `geom_boxplot()` function that is built into ggplot. 

```{r}
# making a box and whisker plot with ggplot.  Here, instead of specifiying point, you specify the geom_boxplot to add the boxplot as another layer
ggplot(compensation, aes(x = Grazing, y = Fruit)) +
geom_boxplot() +
xlab("Grazing treatment") +
ylab("Fruit Production") +
theme_bw()
```
When creating a boxplot, the boxplot is added as another layer on top of the graph.  That being said, we can also add another layer with raw data by incorporating the `geom_point()` function into the code using the `+` sign.

```{r}
ggplot(compensation, aes(x = Grazing, y = Fruit)) +
geom_boxplot() +
geom_point(size = 4, colour = 'lightgrey', alpha = 0.5) +
xlab("Grazing treatment") +
ylab("Fruit Production") +
theme_bw()
```
We can use these plots to analyze the data.  Based on these plots, it appears that fruit production is higher in plants with wider root diameters.  Additionally, grazed plants produce more fruit. 

## Making Histograms of Numeric Variabales
We can use `ggplot()` to make histograms using the `geom_histogram()` function.  That being said, we do have to think about the inputs and how we want to sort or bin them.  This means that the `aes()` function will only have the x-variable. 

```{r}
# creating a histogram using ggplot()
ggplot(compensation, aes(x = Fruit)) +
geom_histogram()

## ‘stat_bin()‘ using ‘bins = 30‘. Pick better value with
## ‘binwidth‘.
```
This does not look quite as neat as we would like it to.  To change this, we will change the binwidth using the following code:

```{r}
# changing the binwidth of the histogram
ggplot(compensation, aes(x = Fruit)) +
geom_histogram(bins = 10)
ggplot(compensation, aes(x = Fruit)) +
geom_histogram(binwidth = 15)
```
## Facets
Faceting allows us to produce a matrix of graphs that are automatically structured by the factors in your data. The function that we will use to do this is called `facet_wrap()`.  See it in practice below:

```{r}
# using faceting to organize the data, you have to preceed the grouping variable with a "~" symbol for this to work

ggplot(compensation, aes(x = Fruit)) +
geom_histogram(binwidth = 15) +
facet_wrap(~Grazing)
```
# Saving Graphs For Future Use
There are a few different ways that we can save these plots so we can use them in the future.  The first method is to go to the "Plots" tab on the right and go to the "Export" button.

Another method is to use the `ggsave()` function built into `ggplot2`.  It works by saving the figure int the "Plots" window to a name that you are able to specify.  Depending on the file suffix that you use, you can dictate the file type that is created. 

```{r}
# Using ggsave() to save our histogram above
ggsave("ThatCoolHistogramOfFruit.png")
```



