---
title: "Homework 21 - Working With Data in R"
output: html_notebook
---

# Overview
During this homework assignment we are going to take a look at different ways that you can visualize data.  To start, I am going to import the libraries that we are going to need to import and manipulate the data. 

```{r}
# clear the R environment and import the libraries that we need
rm(list = ls())
library(here)
library(ggplot2)
library(dplyr)
```

The data that we are going to be working with is the "compensation.csv" data that we have been working with. To import it, I will use the `read.csv()` and the `here()` functions.

```{r}
# import the data 
compensation <- read.csv(here("Data", "compensation.csv"))

# open the data to ensure that it imported correctly
View(compensation)
```

# Working with `ggplot2()`
The first thing that we are going to do is try and create a simple scatter plot of the data.  We will do this using `ggplot2()`.

```{r}
# Use ggplot to make a simple scatterplot
ggplot(compensation, aes(x = Root, y = Fruit)) + geom_point()
```

